\LARGE \bf Energy-Aware Dynamic Mission Planning Algorithm for UAVs

### 2020-11-25 icra-2021.tex
trying to implement Ulrik's review (still to go)

Many scenarios involving unmanned aerial vehicles (UAVs) such as precision agriculture, search and rescue, and surveillance require high autonomy but have limited energy budgets. A typical example of a mission in these scenarios is a UAV flying a trajectory and performing some on-board computations with little human interaction. For instance, the UAV might detect ground patterns and inform other ground-based actors. Energy-aware dynamic mission planning finds optimal tradeoffs between the trajectory, computations, and energy requirements. However, current generic planning solutions for UAVs flying outdoors do not plan the mission dynamically, nor are energy-aware. They are often semi-autonomous; the mission is static and usually defined using mission planning software~\cite{daponte2019review}. Such a state of practice has prompted us to propose an \emph{energy-aware dynamic mission planning algorithm} for UAVs. The algorithm combines and generalizes some of the past body of knowledge on mobile robot planning problems and addresses the increasing \emph{computational demands} and their relation to energy consumption, path, and autonomy for the UAV mission planning problem.

Planning algorithms literature for mobile robots is by now large and related to topics such as trajectory generation and path planning. Generally, the algorithms select an energy-optimized trajectory~\cite{mei2004energy}, e.g., by maximizing the operational time~\cite{wahab2015energy}. However, they apply to a small number of robots~\cite{kim2005energy}, and focus on planning the trajectory for these robots~\cite{kim2008minimum}, despite compelling evidence for the energy consumption also being influenced by computations~\cite{mei2005case}. Given the availability of powerful GPU-equipped mobile hardware, the use of computations is expected to increase in the near future. Mission planning algorithms, which include a broader concept of a mission being a set of tasks along with a motion plan, all focus on the trajectory~\cite{mei2005case,mei2006deployment} and apply to a small number of robots~\cite{sadrpour2013mission,sadrpour2013experimental}. For UAVs specifically, rotorcrafts have gained interest in terms of algorithms for energy-optimized trajectory generation~\cite{morbidi2016minimum,kreciglowa2017energy}.

Our goal is to complete the mission with the highest possible computations and adjustments as the UAV flies and its batteries drain. 
The algorithm plans both tasks and trajectory unlike most of the past planning algorithms literature. It guides the UAV using a vector field~\cite{de2017guidance} that converges smoothly to the planned trajectory. The vector field for planning is widely discussed in the literature~\cite{lindemann2005smoothly,gonccalves2010vector,panagou2014motion,zhou2014vector,kapitanyuk2017guiding}. 
The user specifies an initial mission plan--the trajectory and the tasks--along with an energy budget and some parameters: the Quality of Service (QoS) computations are relative to the tasks. The trajectory-explicit equations (TEEs) adjustments are relative to the path. The computations specify a given task's computational level. The adjustments are a way to alter the TEE. The algorithm replans the initial mission plan as the energy budget changes due to atmospheric interferences.
In the remainder of the paper, we adopt the following notation. We refer to the values of mission-specific QoS and TEEs parameters as computations and adjustments, we refer to the constraints sets that delimit such computations and adjustments as QoS and TEEs sets, and we refer to the current trajectory as TEE.

The algorithm contributes to the planning literature further with a periodic energy model that accounts for the uncertainty (atmospheric interference). Mission periodicity is usually present due to the mission's repetitive patterns~\cite{seewald2020mechanical}. Indeed UAV scenarios often iterate over a set of tasks and trajectories (e.g., monitoring or search and rescue). The mission is periodic, so we expect the energy to approximately evolve periodically. The observation applies broadly; repetitive patterns have been observed in scenarios as far as space rovers~\cite{seewald2020beyond}. The algorithm addresses the periodicity through Fourier analysis. The uncertainty with a state estimator. It outputs the control input (adjustments and computations) using output model predictive control (MPC).

@@

\begin{lem}[Signals equality]\label{lem:eqv}Suppose there is no state and output uncertainty (the contributions $\mathbf{w}\in\mathbb{R}^m,v\in\mathbb{R}$ are zero) and the control is zero. Suppose further matrices $A,C$ are described by Equation~(\ref{eq:state-details}), and the initial guess $\mathbf{q}_0$ is $\mathbf{q}_0=\begin{bmatrix}a_0 & a_1/2 & b_1/2 & \cdots & a_r/2 & b_r/2\end{bmatrix}$. Then Equation~(\ref{eq:fourier}) describes the same signal as Equation~(\ref{eq:state-perf}).
\end{lem}
\begin{proof}
The equivalence of the models is trivial and the coefficients are chosen such that the two Equations produce the same expression. We refer the reader to Appendix~\ref{app:proof-eqv} for a formal proof.
\end{proof}

@@

Let us consider the second expression. It represents the linear combination of all the coefficients of the state at time $t$. It can also be expressed in the following form
\begin{equation}\label{eq:proof-output}\begin{split}
  \gamma_0 e^{\lambda_0t}v_0/T+&(1/T)\sum_{j=1}^r{\gamma_j e^{\lambda_j^0t}v_j^0}+\\&(1/T)\sum_{j=1}^r{\gamma_j e^{\lambda_j^1t}v_j^1}.
\end{split}\end{equation}

@@

The initial guess is build such that the sum of the coefficients is the same in both the signals. The output matrix is built as follows. The coefficient $1/T$ accounts for the period in Equations~(\ref{eq:proof-complex}--\ref{eq:proof-output}) and~(\ref{eq:fourier}). At time instant zero, the coefficients $b_j$ are not present (and the coefficients $a_j$ are doubled for each $j=1,2,\dots,r$). Thus to match the outputs $h(t)=y(t)$ we define $C=1/T\begin{bmatrix}1 & 1 & 0 & \cdots & 1 & 0\end{bmatrix}$. We thus conclude that the lemma holds.

We note that the signal would still be periodic with another linear combination of coefficients (for instance, $C=\begin{bmatrix}1 & 0 & 1 & \cdots & 0 & 1\end{bmatrix}$, or $\begin{bmatrix}1 & \cdots & 1\end{bmatrix}$). Even if both the signals $h$ and $y$ would be still periodic, they would not be equal.

>>>>

Many scenarios involving unmanned aerial vehicles (UAVs), such as precision agriculture, search and rescue, and surveillance, require high autonomy but have limited energy budgets. A typical example of a mission in these scenarios is a UAV flying a trajectory and performing some on-board computations with little human interaction. For instance, the UAV might detect ground patterns and inform other ground-based actors. We are interested in energy-aware dynamic mission planning, which would find optimal tradeoffs between the trajectory, computations, and energy requirements. Current generic planning solutions for outdoor UAVs do not plan the mission dynamically, nor are energy-aware. They are often semi-autonomous; the mission is static and usually defined using mission planning software~\cite{daponte2019review}. Such a state of practice has prompted us to propose an \emph{energy-aware dynamic mission planning algorithm} for UAVs. The algorithm combines and generalizes some of the past body of knowledge on mobile robot planning problems and addresses the increasing \emph{computational demands} and their relation to energy consumption, path, and autonomy for the UAV mission planning problem.

Planning algorithms literature for mobile robots includes topics such as trajectory generation and path planning. Generally, the algorithms select an energy-optimized trajectory~\cite{mei2004energy}, e.g., by maximizing the operational time~\cite{wahab2015energy}. However, they apply to a small number of robots~\cite{kim2005energy}, and focus exclusively on planning the trajectory for these robots~\cite{kim2008minimum}, despite compelling evidence for the energy consumption also being significantly influenced by computations~\cite{mei2005case}. Given the availability of powerful GPU-equipped mobile hardware, the use of computations is expected to increase in the near future. Mission planning algorithms, which include a broader concept of a mission being a set of tasks along with a motion plan, all focus on the trajectory~\cite{mei2005case,mei2006deployment} and apply to a small number of robots~\cite{sadrpour2013mission,sadrpour2013experimental}. For UAVs specifically, rotorcrafts have gained interest in terms of algorithms for energy-optimized trajectory generation~\cite{morbidi2016minimum,kreciglowa2017energy}.

Unlike most of the past planning algorithms literature, our algorithm plans both the computations and the trajectory. We guide the UAV using a vector field~\cite{de2017guidance} that converges smoothly to the planned trajectory. The use of vector fields for planning is widely discussed in the literature~\cite{lindemann2005smoothly,gonccalves2010vector,panagou2014motion,zhou2014vector,kapitanyuk2017guiding}. To express the trajectory, we use a mathematical function that we refer to as the trajectory explicit equation (TEE). We plan the computations of the tasks running on the UAV using a modeling tool presented in our previous work~\cite{seewald2019coarse} and explore the trade-offs with the trajectory. In our approach, the user specifies an initial mission plan--the TEEs, the energy budget, and some parameters. The parameters are the maximum, minimum, and current levels of the computations and trajectory. The energy budget is the battery capacity and some other battery parameters. The algorithm replans the initial mission plan as the remaining battery state of charge (SoC) unexpectedly changes due to, e.g., atmospheric interferences. Our goal is to complete the mission with the highest possible current computations and trajectory parameters as the UAV flies and its batteries drain. In the remainder of the paper, we adopt the following notation. We refer to the current computations and trajectory parameters as computations and trajectory parameters. We refer to just parameters when we deal with both the computations and trajectory parameters. We refer to the constraints sets that delimit the maximum and minimum level of parameters as computations sets and trajectory sets. There can be multiple or none computations sets per each task, and the same applies to the trajectory. The term parameters is not to be confused with the battery parameters, which we always mention explicitly. These are fixed values that are not replanned by the algorithm.

% Ulrik, I don't agree with a main. The main contibution is the inclusion of computations and trajectory in the planning. Should I stress that? The periodicity helps us to achieve that.
To achieve the dynamic mission planning, we further introduce and formally proof a periodic energy model that accounts for the uncertainty (atmospheric interference). Mission periodicity is often present due to the mission's repetitive patterns~\cite{seewald2020mechanical}. Indeed UAV scenarios often iterate over a set of tasks and trajectories (e.g., monitoring or search and rescue). Given that the mission is periodic, we expect the energy to approximately evolve periodically. % Adam you need to fix from here on xD
The observation applies broadly; repetitive patterns have been observed in scenarios as far as space rovers~\cite{seewald2020beyond}. The algorithm addresses the periodicity through Fourier analysis. The uncertainty with a state estimator. It outputs the control input (adjustments and computations) using output model predictive control (MPC).

@@

\begin{lem}[Signals equality]\label{lem:eqv}Suppose there is no state and output uncertainty (the contributions $\mathbf{w}\in\mathbb{R}^m,v\in\mathbb{R}$ are zero) and the control is zero. Suppose further matrices $A,C$ are described by Equation~(\ref{eq:state-details}), and the initial guess $\mathbf{q}_0$ is $\mathbf{q}_0=\begin{bmatrix}a_0 & a_1/2 & b_1/2 & \cdots & a_r/2 & b_r/2\end{bmatrix}$. Then Equation~(\ref{eq:fourier}) describes the same signal as Equation~(\ref{eq:state-perf}).
\end{lem}
\begin{proof}
The equivalence of the models is trivial and the equality of the two signals is achieved by a proper choice of items of matrices $A,C$ and the initial guess $\mathbf{q}_0$. We refer the reader to Appendix~\ref{app:proof-eqv} for a formal proof the equality.
\end{proof}

@@

Let us consider the second expression. It represents the linear combination of all the coefficients of the state at time $t$. It can also be expressed in the following form
\begin{equation}\label{eq:proof-output}\begin{split}
  F\mathbf{q}(t)/T=\gamma_0 e^{\lambda_0t}v_0/T+&(1/T)\sum_{j=1}^r{\gamma_j e^{\lambda_j^0t}v_j^0}+\\&(1/T)\sum_{j=1}^r{\gamma_j e^{\lambda_j^1t}v_j^1}.
\end{split}\end{equation}

@@

The initial guess is build such that the sum of the coefficients is the same in both the signals. The output matrix is built as follows. The coefficient $1/T$ accounts for the period in Equations~(\ref{eq:proof-complex}--\ref{eq:proof-output}) and~(\ref{eq:fourier}). At time instant zero, the coefficients $b_j$ are not present (and the coefficients $a_j$ are doubled for each $j=1,2,\dots,r$). To match the outputs $h(t)=y(t)$--or equivalently $F\mathbf{q}(t)/T=C\mathbf{q}(t)$--we define $C=1/T\begin{bmatrix}1 & 1 & 0 & \cdots & 1 & 0\end{bmatrix}$. We thus conclude that the lemma holds.

We note that the signal would still be periodic with another linear combination of coefficients (for instance, $C=c\begin{bmatrix}1 & 0 & 1 & \cdots & 0 & 1\end{bmatrix}$, or $c\begin{bmatrix}1 & \cdots & 1\end{bmatrix}$ for a constant $c\in\mathbb{R}$). Even if both the signals $h$ and $y$ would be still periodic, they would not be equal.
